---
title: "Getting Started with randomforge"
author: "Friedrich Pahlke"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with randomforge}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

`randomforge` is an open-source R package providing a transparent and modular 
framework for clinical trial randomization.  
This vignette gives a quick introduction on how to install the package, create a
simple project, configure a permuted block randomization method, and generate 
subject allocations.

If you are new to randomization frameworks or want a minimal working example, 
this is a good place to start.


# Installation

At this stage, `randomforge` is available only on GitHub:

```{r}
#| eval: FALSE
# install.packages("remotes")
remotes::install_github("RCONIS/randomforge")
```

# Creating a Randomization Project

Every workflow begins with a `RandomProject` stored inside an in-memory 
`RandomDataBase`:

```{r}
library(randomforge)

# Create an in-memory randomization database
randomDataBase <- getRandomDataBase()

# Define a project
randomProject <- getRandomProject("Example Trial")

# Store the project in the database
randomDataBase$persist(randomProject)
```

A project groups all configurations, subjects, and resulting allocations.


# Defining a Randomization Configuration

A configuration defines:

- treatment arms  
- allocation parameters  
- random number buffer settings  
- seeds  
- optional stratification  

Example:

```{r}
config <- getRandomConfiguration(
    randomProject        = randomProject,
    treatmentArmIds      = c("A", "B"),
    seed                 = createSeed(),
    ravBufferMinimumSize = 1000L,
    ravBufferMaximumSize = 10000L
)
config

# Store the configuration
randomDataBase$persist(config)
```


# Creating a Block Randomization Method

`randomforge` currently supports **permuted block randomization** (PBR) as a 
fully working implementation.

You can define variable block sizes:

```{r}
# Define variable block sizes
blockSizes <- getBlockSizes(config$treatmentArmIds, c(4, 6))

# Define a block randomization method
blockSizeRandomizer <- getRandomBlockSizeRandomizer(blockSizes)
blockSizeRandomizer

randomMethodPBR <- getRandomMethodPBR(
    blockSizes              = blockSizes,
    fixedBlockDesignEnabled = FALSE,
    blockSizeRandomizer     = blockSizeRandomizer
)
```

# Create a Random Allocation Value Service

```{r}
# Create a random allocation value service
ravService <- getRandomAllocationValueService()
ravService$createNewRandomAllocationValues(config)
```

Quality control of the random numbers used for randomization can be visualized by

```{r}
ravService |>
    plot(usedValuesOnly = FALSE)
```

Other key performance indicators are planned for future releases.

# Running Randomization

To generate assignments, create a random allocation value service, then call 
`getNextRandomResult()`.

```{r}
# Create a few randomization results
resultList <- lapply(1:12, function(i) {
    getNextRandomResult(
        randomDataBase               = randomDataBase,
        randomProject                = randomProject,
        randomMethod                 = randomMethodPBR,
        randomAllocationValueService = ravService
    ) |> suppressMessages()
})
```

```{r}
ravService |>
    plot()
```

# Inspecting the Results

All subjects and allocations stored in the database can be displayed as a data
frame:

```{r}
#| eval: false
# Convert results to a data frame
resultData <- randomDataBase |>
    as.data.frame()
resultData
```

```{r}
#| echo: false
resultData <- randomDataBase |>
    as.data.frame()
resultData |>
    knitr::kable()
```

# Exporting to Excel (Optional)

`randomforge` supports exporting subject lists or randomization results via 
`writeExcelFile()`:

```{r}
#| eval: false
writeExcelFile(resultData, "randomization_list.xlsx")
```


# What’s Next?

The project is in an early phase, and many extensions are planned:

- covariate-adaptive methods (e.g., minimization)
- response-adaptive techniques  
- stratified and center-based workflows  
- audit trail and reporting components  
- Shiny and API integration  

To learn how to contribute, see the vignette:

**`vignette("randomforge_contribution")`**


# Thank You

We hope this vignette helps you get started with `randomforge`.  
Feedback and suggestions are very welcome via GitHub issues.
